---
title: "InstaCart Market Basket Analysis"
author: "Prashant Nair"
date: "01 March 2018"
output:
  html_document:
    fig_height: 4
    fig_width: 7
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###                                AFFINITY ANALYSIS 

### Understanding Market Basket Analysis / Assosiation rule mining 

MBA allows us to identify items that are frequently bought together.Typically the output of an MBA is in  the form of rules. The rules can be simple {A ==> B}, when a customer buys item A then it is (very) likely that the customer buys item B. More complex rules are also possible {A, B ==> D, F}, when a customer buys items A and B then it is likely that he buys items D and F.

### Read in the data  
###  Note : Input in some sample data in case the processing and memory is low 

```{r message=FALSE, warning=FALSE, results='hide'}
library(data.table)
library(dplyr)
library(ggplot2)
library(knitr)
library(stringr)
library(DT)
library(arules)
library(arulesViz)
library(plotly)
library(visNetwork)
library(igraph)

aisles <- fread('./dataForR/aisles.csv')
departments <- fread('./dataForR/departments.csv')
order_products_prior <- fread('./dataForR/order_products__prior.csv',nrows=20000) #Needed
order_products_train <- fread('./dataForR/order_products__train.csv')
orders <- fread('./dataForR/orders.csv')
products <- fread('./dataForR/products.csv')    # Needed 
```

### Data Preparation , Creating a customers shopping basket 

```{r, message=FALSE,warning=FALSE,}
# get the shopping baskets
order_baskets <- order_products_prior %>% 
  inner_join(products, by="product_id") %>% 
  group_by(order_id) %>%
  summarise(basket = as.vector(list(product_name)))

# compute transactions
transactions <- as(order_baskets$basket, "transactions")
```
### Frequent Itemsets 
```{r, fig.height=8}
###   Frequent Item Sets #####
support <- 0.008
itemsets <- apriori(transactions, parameter = list(target = "frequent itemsets", supp=support, minlen=2), control = list(verbose = FALSE))

par(mar=c(5,18,2,2)+.1)
sets_order_supp <- DATAFRAME(sort(itemsets, by="support", decreasing = F))
barplot(sets_order_supp$support, names.arg=sets_order_supp$items, xlim=c(0,0.02), horiz = T, las = 2, cex.names = .8, main = "Frequent Itemsets", col="blue")
mtext(paste("support:",support), padj = .8)
```

### Association Rules

### Setting the Support and Confidence Intervals.

### What is support?

A value of Support =0.02 means that an item will be considered as frequent if at 
least 2 percent of all the baskets contain it. 

### What is Confidence?

confidence is a measure of the strength of an association rule. It is the frequency of occurrence of the right-hand items in the rule from among those baskets that contain the items on the left-hand side of the rule.

### Rule1 : 
using  a low support threshold to generate strong rules even for items that are less frequent

```{r, message=FALSE, warning=FALSE}
#3006775 rules

rules1 <- apriori(transactions, parameter = list(supp = 0.0000001, conf = 0.01, maxlen=3), control = list(verbose = FALSE))

plotly_arules(rules1)
```
### Rule 2 : 

```{r, message=FALSE, warning=FALSE}
rules <- apriori(transactions, parameter = list(supp =  0.003269976, conf = 0.01, maxlen=3), control = list(verbose = FALSE))

plotly_arules(rules)
```

### Rule 3: 

```{r, message=FALSE, warning=FALSE}
#PLOT 4
rules2 <- apriori(transactions, parameter = list(supp = 0.001, conf = 0.4, maxlen=3), control = list(verbose = FALSE))
inspect(sort(rules2, by="lift")[1:10])
```


## Rule 4 : High Confidence and less support 

```{r, message=FALSE, warning=FALSE}
#PLOT 5
rules3 <- apriori(transactions, parameter = list(supp = 0.005, conf = 0.1, maxlen=3), control = list(verbose = FALSE))
inspect(sort(rules3, by="lift")[1:10])

```


### Graph Analysis for rules  

### The rules obtained from association mining form a network graph.
Node : Each product in the customer shopping 
Edge : each rule "from ==> to" is an edge of the graph


```{r, message=FALSE}
# uses rules2 top 10 by lift PLOT 4
subrules2 <- head(sort(rules3, by="confidence"),10)
ig <- plot( subrules2, method="graph", control=list(type="items") )

ig_df <- get.data.frame( ig, what = "both" )
nodesv <- data.frame(
  id = ig_df$vertices$name
  ,value = ig_df$vertices$lift # could change to lift or confidence
  ,title = ifelse(ig_df$vertices$label == "",ig_df$vertices$name, ig_df$vertices$label)
  ,ig_df$vertices
) 

edgesV = ig_df$edges

```

### Network graph visualization Part 2

```{r, message=FALSE}
visNetwork(nodes = nodesv, edges = edgesV, width = 900, height = 700) %>%
  visNodes(size = 10) %>%
  visLegend() %>%
  visEdges(smooth = FALSE) %>%
  visOptions(highlightNearest = TRUE, nodesIdSelection = TRUE) %>%
  visInteraction(navigationButtons = TRUE) %>%
  visEdges(arrows = 'from') %>%
  visPhysics(
    solver = "barnesHut",
    maxVelocity = 35,
    forceAtlas2Based = list(gravitationalConstant = -6000)
  )
```



